Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828

### 2016-07-12 inst/examples/20160715-slides-Krakow.Rmd
ok

---
title: "Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828"
author: "Mikko Tolonen, Jani Marjanen, Hege Roivainen, Leo Lahti"
date: "`r Sys.Date()`"
output: 
  beamer_presentation
---

@@

library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)
library(devtools)
# install_github("ropengov/gisfin")
# devtools::install_github('cttobin/ggthemr')
library(gisfin)
#library(ggthemr)

@@

sivu 3. Datamäärät koodista tähän takaisin. Deletoin tän vahingossa aikaisemmin kun koodi ei toiminut.

sivu 8. Karttakuvassa Greifswald, Lund ei näy. Tätä olisi hyvä saada kehitettyä, tällaisena tää ei ihan toimi.

@@

## From particular data to global perspective?

**Idea**: to study early modern knowledge production, all of it (1470-1830)

**How?** By combining library catalogues (Fennica, Kungliga, ESTC, CERL's HPB database)

**Important**: to clean up the data (90% of the work!) and to take into account how particular data collections (Finnish National Bibliography, English Short-Title Catalogue etc.) have been formed.

@@

**Overcoming a national gaze** by focusing on how knowledge production happened through a more complex network of cities.

@@

### Data (1640-1828):

TÄHÄN DATA MÄÄRÄ KOODISTA!

### Method

- Open source tools for library catalogues (Github/Fennica)

- Fully transparent and reproducible

- Scales up to new catalogues and tasks

- Benefit no. 1: once the work has been done, the data is available for everyone to use and to enrich. The generic data analytical ecosystem is designed with even unforeseen changes in mind.

@@

### Outline of Quantitative Study of Finnish Knowledge Production, 1640-1828

@@

--> Visible roots of the diversity of "Finnishness" reflected in knowledge production, 1640-1828.

@@

### Turku as a Swedish University Town among Uppsala, Tartu and Stockholm (+ Greifswald and Lund)

```{r map1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=5}
library(dplyr)
library(ggmap)
library(ggplot2)
library(gisfin)
source("funcs.R")
theme_set(get_theme_map())
mapdata <- filter(df0,
          (catalog == "Kungliga" &
	      publication_place %in% c("Stockholm", "Uppsala",
	                         "Greifswald", "Lund", "Tartu"))|
	  (catalog == "Fennica" &
	      publication_place == "Turku")) %>%
      group_by(publication_decade, publication_place,
	       latitude, longitude) %>%
      summarise(n = n())

@@

### Fennica drop in 1700-1721, Kungliga peak in 1760-1770s, effect of Napoleonic wars (1803-1815)

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4, out.width="200px"}
df2 <- df0 %>% group_by(publication_decade, catalog) %>%
               summarise(n = n())
df2$highlight <- rep(FALSE, nrow(df2))
df2$highlight[df2$catalog == "Kungliga" &
	      df2$publication_decade %in% c(1770, 1810)] <- TRUE
df2$highlight[df2$catalog == "Fennica" &
	      df2$publication_decade %in% c(1710, 1810)] <- TRUE
v <- seq(1500, 1800, 50)
theme_set(theme_bw(20))
p <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = catalog), color = "black") +
     geom_point(aes(shape = catalog, color = highlight, size = highlight)) +
     ylab("Title count (n)") + xlab("Publication year") +
     scale_color_manual(values = c("darkgray", "black")) +
     theme(legend.position = "none") +
     scale_x_continuous(breaks = v, labels = v)
     # ggtitle("Overall publishing activity")
print(p)
```

@@

```{r publishingactivitycomparisons, echo=FALSE, message=FALSE, cache=TRUE, fig.width=6, fig.height=5, fig.show="hold", out.width="170px"}
v <- seq(1500, 1800, 50)
selected.places = c("Turku", "Stockholm")
df2 <- df0 %>% group_by(publication_decade, publication_place, catalog) %>%
               summarise(n = n()) %>%
	       filter(publication_place %in% selected.places) 
df2$catalog = factor(df2$catalog, levels = rev(c("Fennica", "Kungliga")))
df2$publication_place = droplevels(factor(df2$publication_place))
df3 = as.data.frame(spread(df2, publication_decade, n, fill = 0))
df2 = melt(df3)
colnames(df2) = c("publication_place", "catalog", "publication_decade", "n")
df2$publication_decade = as.numeric(as.character(df2$publication_decade))
df2$n = as.numeric(as.character(df2$n))
for (catal in unique(df2$catalog)) {
  p <- ggplot(subset(df2, catalog == catal), aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = publication_place)) +
     geom_point(aes(shape = publication_place), size = 3) +
     ylab("Title count (n)") + xlab("Publication year") +  
     ggtitle(catal) +
     scale_x_continuous(breaks = v, labels = v)     
  print(p)

@@

TODO: Vaihdettiin tähän Stockholm, riittää julkaisumäärät. Miksi katalogina lukee Fennica? Pitää olla Kungligan mukaan data tässä.

@@

@@

Extra Slide:

### Duplicate publications

```{r duplicates, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=5, fig.width=10}
# DUPLICATE FRACTION
df <- df.combined.preprocessed %>%
        filter(publication_year >= 1640 & publication_year <=1828) %>%
        group_by(catalog, publication_decade) %>%
	summarise(duplicates = 100 * mean(remove))

p <- ggplot(df, aes(x = publication_decade, y = duplicates, fill = catalog)) +
       geom_bar(position = "dodge", stat = "identity") +
       ylab("Duplicates (%)") +
       scale_fill_manual(values = c("blue", "red")) + 
       ggtitle(paste("Certain duplicates Fennica-Kungliga", paste(range(na.omit(df$publication_decade)), collapse = "-")))

>>>>

---
title: "Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828"
author: "Mikko Tolonen, Jani Marjanen, Hege Roivainen, Leo Lahti"
date: "`r Sys.Date()`"
output: 
  beamer_presentation:
    theme: "default"
    colortheme: "dolphin"
---

@@

library(ggmap)
library(gisfin)
library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)
library(devtools)
# install_github("ropengov/gisfin")
# devtools::install_github('cttobin/ggthemr')
library(gisfin)
#library(ggthemr)

@@

@@

## Global knowledge production and large-scale library catalogues

**Idea**: Study early modern knowledge production, all of it (1470-1830)

**How?** Combine library catalogues (Fennica, Kungliga, ESTC, CERL HPB database)

**Important**: cleaning up the data (90% of the effort!) and considering how particular data collections (Finnish National Bibliography, English Short-Title Catalogue etc.) have been formed.

@@

**Overcoming a national gaze** focus on knowledge production through more complex cross-national networks

**Data: bibliographic catalogues 1640-1828**

- Fennica: `r nrow(subset(df0, catalog == "Fennica" & publication_year >= 1640 & publication_year < 1828))` documents
- Kungliga: `r nrow(subset(df0, catalog == "Kungliga" & publication_year >= 1640 & publication_year < 1828))` documents

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=7}
library(dplyr)
library(ggplot2)
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>% group_by(catalog, publication_decade) %>% tally()
p <- ggplot(df, aes(x = publication_decade, y = n, color = catalog)) + 
       geom_point() + 
       geom_line() + 
       ylab("Documents (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen"))
print(p)
```

@@

### Methods: Open data analytical ecosystem

- **Automated** scales well with new data and unforeseen changes

- **Reusable** the polished data can be reused and further enriched

- **Customized** dedicated R packages for bibliographic analysis combine automation with flexibility

- **Open source** fully transparent and reproducible data analytics  
  https://github.com/rOpenGov/fennica

![Data Science Workflow](data-science.png)  
\tiny{(Figure by Hadley Wickham)}

@@

### Examples of the tools

- **Parsers** From original MARC/XML to CSV to RData

- **Preprocessing** Cleaning up the fields.

```{r}
library(bibliographica)
polish_physical_extent("iii-xxiv, 118, [2] p.")$pagecount
polish_place("Sur l'imprimea londres")
```

**Analysis**

```{r}
library(knitr)
kable(as.data.frame(top(df0, "publication_place", 3)))
```

**Visualization**

```{r, fig.width=10, fig.height=7}
titlecount_timeline(subset(df0, catalog == "Fennica"), 
                         "publisher", nmin = 500, mode = "n")$plot
```

### Outline

**Quantitative Study of Finnish Knowledge Production, 1640-1828**

@@

Visible roots of the diversity of "Finnishness" reflected in knowledge production, 1640-1828.

@@

### Turku as a Swedish University Town among Uppsala, Tartu and Stockholm

TODO Karttakuvassa Greifswald, Lund ei näy. Tätä hyvä saada kehitettyä, tällaisena ei toimi. LL: ei näy sen takia että näytettynä on vuosikymmen 1650 jolloin Lundissa on 0 dokumenttia ja Greifswaldissa 1. Muissa kymmeniä tai satoja. Data näkyy kuten pitääkin.

```{r map1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=5, eval=FALSE}
library(dplyr)
library(ggmap)
library(ggplot2)
library(gisfin)
source("funcs.R")
theme_set(get_theme_map())
mapdata <- filter(df0,
          (catalog == "Kungliga" &
	      publication_place %in% c("Stockholm", "Uppsala",
	                         "Greifswald", "Lund", "Tartu"))|
	  (catalog == "Fennica" &
	      publication_place == "Turku")) %>%
      group_by(publication_decade, publication_place,
	       latitude, longitude) %>%
      summarise(n = n())

@@

### Publishing activity over time

- 1700-1721 Fennica drop
- 1760-1770s Kungliga peak 
- 1803-1815 Napoleonic wars

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4, out.width="200px"}
library(magrittr)
library(ggplot2)
library(dplyr)
df2 <- df0 %>% group_by(publication_decade, catalog) %>%
               summarise(n = n())
df2$highlight <- rep(FALSE, nrow(df2))
df2$highlight[df2$catalog == "Kungliga" &
	      df2$publication_decade %in% c(1770, 1810)] <- TRUE
df2$highlight[df2$catalog == "Fennica" &
	      df2$publication_decade %in% c(1710, 1810)] <- TRUE
v <- seq(1500, 1800, 50)
theme_set(theme_bw(20))
p <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = catalog), color = "black") +
     geom_point(aes(shape = catalog, color = highlight, size = highlight)) +
     ylab("Title count (n)") + xlab("Publication year") +
     scale_color_manual(values = c("darkgray", "black")) +
     theme(legend.position = "none") +
     scale_x_continuous(breaks = v, labels = v)
     # ggtitle("Overall publishing activity")
print(p)
```

@@

```{r publishingactivitycomparisons, echo=FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=5, fig.show="hold", out.width="150px"}
v <- seq(1500, 1800, 100)
selected.places = c("Turku", "Stockholm")
df2 <- df0 %>% group_by(publication_decade, publication_place, catalog) %>%
               summarise(n = n()) %>%
	       filter(publication_place %in% selected.places) 
df2$catalog = factor(df2$catalog, levels = rev(c("Fennica", "Kungliga")))
df2$publication_place = droplevels(factor(df2$publication_place))
df3 = as.data.frame(spread(df2, publication_decade, n, fill = 0))
df2 = melt(df3)
colnames(df2) = c("publication_place", "catalog", "publication_decade", "n")
df2$publication_decade = as.numeric(as.character(df2$publication_decade))
df2$n = as.numeric(as.character(df2$n))
for (catal in unique(df2$catalog)) {
  p <- ggplot(subset(df2, catalog == catal), aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = publication_place)) +
     geom_point(aes(shape = publication_place), size = 3) +
     ylab("Title count (n)") + xlab("Publication year") +  
     ggtitle(catal) +
     scale_x_continuous(breaks = v, labels = v) + 
     theme(axis.text.x = element_text(angle = 0))
  print(p)

@@

TODO: Vaihdettiin tähän Stockholm, riittää julkaisumäärät.

@@

Source code for these slides:
- https://github.com/rOpenGov/fennica/blob/master/inst/examples/20160715-slides-Krakow.Rmd

@@

### Duplicate publications in Fennica/Kungliga

```{r duplicates, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=5, fig.width=10}
df <- df.combined.preprocessed %>%
        filter(publication_year >= 1640 & publication_year <=1828) %>%
        group_by(catalog, publication_decade) %>%
	summarise(duplicates = 100 * mean(remove))

p <- ggplot(df, aes(x = publication_decade, y = duplicates, fill = catalog)) +
       geom_bar(position = "dodge", stat = "identity") +
       ylab("Duplicates (%)") +
       scale_fill_manual(values = c("blue", "red")) 
       #ggtitle(paste("Certain duplicates Fennica-Kungliga", paste(range(na.omit(df$publication_decade)), collapse = "-")))
